### 2014-05-12 standing-variation-paper.tex
edited text through the probability of parallel mig

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{natbib}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{latexml} % for \iflatexml

@@

For now we ignore geographic hetergeneity, and assume that the species range
is a homogenous, one- or two-dimensional region. 
There are two selective classes -- the {\em neutral} type, and the {\em mutated} type.
We assume that different mutations are distinguishable --
either as selectively equivalent mutations, or by linked variation.
For a long time in the past, the mutated type has been at a selective disadvantage (long enough to be at selection-mutation equilibrium),
but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
After fixation, alleles are either descended
from families of mutants present as standing variation when the selective regime changed,
or from new mutants arising since that time.

For concreteness, suppose that before time $t=0$,
the mutant type has fitness $1-s_d$ relative to the neutral type
(i.e.\ it produces on average $1-s_d$ times the number of offspring per generation),
and that after time $t=0$,
the mutant type has fitness $1+s_b$,
where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
(If the mutation is recessive, XXX more complicated things happen.)
As for the other parameters,
suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
and that the mean squared distance between parent and child is $\sigma^2$ (the {\em dispersal variance}).
The species occupies an area with mean density
$\rho$ haploid individuals (or chromosomes) per unit area.
\marginnote{XXX or do diploid? below is haploid.}

@@

XXX More discussion on whether this is a good approximation?  
Compute typical size and radius of a standing cluster?

recall PPP computations; draw cones of opportunity for new mutations

@@

In \citet{ralphcoop2010}, studying the model without standing variation,
we defined a {\em characteristic length} which gave the spatial scale across with mutants with distinct origins would establish.
This was proportional to the mean distance between neighboring established mutants,
but had the advantage of being easier to calculate.
Furthermore, the time scale over which adaptation occurred could be found by dividing the characteristic length 
by the speed at which the mutants spread.
We will define the characteristic length for this model,
as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)

@@

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{natbib}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{hyperref}
%\usepackage{latexml} % for \iflatexml
\usepackage{color}
\newcommand{\gc}[1]{{\it\color{green}(#1)} }
\newcommand{\plr}[1]{{\it\color{blue}(#1)}}

@@

We assume that the species range
is a homogenous, one- or two-dimensional region. 
There are two selective classes -- the {\em neutral} type, and the {\em mutated} type.
We assume that different mutations are distinguishable --
either as selectively equivalent mutations, or by linked variation.
For a long time in the past, the mutated type has been at a selective disadvantage (long enough to be at selection-mutation equilibrium),
but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
After fixation, alleles are either descended
from families of mutants present as standing variation when the selective regime changed,
or from new mutants arising since that time.

For concreteness, suppose that before time $t=0$,
the mutant type has fitness $1-s_d$ relative to the neutral type
(i.e.\ it produces on average $1-s_d$ times the number of offspring per generation),
and that after time $t=0$,
the mutant type has fitness $1+s_b$,
where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
We assume that diploid fitness is additive, or at least that the
important early dynamics are determined by the haploid fitness
\gc{heterozygous fintess, with no reference to the fitness of the
  homozygote for the mutant alleles}.
The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
(If the mutation is recessive, XXX more complicated things happen.)
As for the other parameters,
suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
and that the mean squared distance between parent and child is $\sigma^2$ (the {\em dispersal variance}).
The species occupies an area with mean density
$\rho$ haploid individuals (or chromosomes) per unit area.
\marginnote{XXX or do diploid? below is haploid.}

@@

\plr{XXX More discussion on whether this is a good approximation?  
Compute typical size and radius of a standing cluster?

recall PPP computations; draw cones of opportunity for new mutations }

\gc{Point out connection to standard mutation selection balance formula}

@@

In \citet{ralphcoop2010}, studying the model without standing variation,
we defined a {\em characteristic length} which gave the spatial scale across with mutants with distinct origins would establish.
This was proportional to the mean distance between neighboring established mutants,
but had the advantage of being easier to calculate.
Furthermore, the time scale over which adaptation occurred could be found by dividing the characteristic length 
by the speed at which the mutants spread.
We will define the characteristic length for this model,
as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
%(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)

@@

\section{Applications}

% G6PD frequency across world http://www.google.com/imgres?imgurl=http%3A%2F%2Fwww.plosmedicine.org%2Farticle%2Finfo%3Adoi%2F10.1371%2Fjournal.pmed.1001339.g002%2Flargerimage&imgrefurl=http%3A%2F%2Fwww.plosmedicine.org%2Farticle%2Finfo%253Adoi%252F10.1371%252Fjournal.pmed.1001339&h=2938&w=2381&tbnid=oD_tX3ea8IbLfM%3A&zoom=1&docid=k37axqAZDS7ljM&ei=5rdOU5TVDMqj8QGntICYBg&tbm=isch&ved=0CF8QMygLMAs&iact=rc&uact=3&dur=1447&page=1&start=0&ndsp=15

% Spatial distribution of G6PD alleles:  http://www.malariajournal.com/content/12/1/418
